---
title: "Effectiveness of community-based marine reserves in small-scale fisheries"
rtitle: "Community-based marine reserves"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "jvillasenor@bren.ucsb.edu"
abstract: "Al finalizar revisiones"
keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, Mexico, Social-Ecological Systems"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

<!-- Old Abstract -->
<!-- Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. However, overfishing and unsustainable fishing practices threaten the marine environment and jeopardize the wellbeing of coastal communities. A common approach to protect the environment and recover overexploited stocks is to implement no-take marine reserves (areas where all extractive activities are off-limits). In small-scale fisheries, these are sometimes implemented as community-based reserves, where a group of fishers collectively agree to close an area to fishing. While we know that reductions in fishing effort are followed by a series of ecological benefits (increased biomass, abundance, and species diversity), we do not fully understand how environmental and governance dynamics influence the conservation and fisheries benefits of community-based marine reserves. In this work, we evaluate the ecological outcomes of four reserves established by three coastal communities in temperate and tropical ecosystems of Mexico. By combining causal inference techniques with an operationalization of the social-ecological systems framework, we identify the environmental and social conditions that enable reserve effectiveness. Our results show a strong interaction between environmental variation and community organization, which influences reserve effectiveness. For example, the most effective reserve had strong governance structures accompanied with low environmental variability. Thus, even when a community is well organized (and reserves are well enforced), environmental variation can hinder the benefits of a reserve, and vice versa. Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

options(knitr.table.format = "latex")
```

**Titulos alternativos**

"Management and effectiveness of community-based marine reserves in small-scale fisheries"

"Community-based marine reserves in small-scale fisheries"

"Effectiveness of community-based marine reserves: lessons for their management and implementation" (no me encanta)

Algun otro?


**Last update**: `r Sys.Date()`

\clearpage


```{r load_pkgs}
suppressPackageStartupMessages({
  library(sp)
  library(sf)
  library(ggsn)
  library(tmap)
  library(magrittr)
  library(kableExtra)
  library(broom)
  library(ggrepel)
  library(cowplot)
  library(tidyverse)
})


# Load indicator functions
source(file = here::here("scripts", "abundance.R")) #Calculate abundance per transect
source(file = here::here("scripts", "biomass.R")) # Calculate fish biomass per transect
source(file = here::here("scripts", "trophic.R")) # Calculate mean trophic level per transect

# Load analysis functions
source(file = here::here("scripts", "model.R")) #fits the model
source(file = here::here("scripts", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
source(file = here::here("scripts", "robust_se_list.R")) #heteroskedastic-robust Standard Errors for stargazer
source(file = here::here("scripts", "plot_model.R")) #plots effect sizes
```

# Introduction

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. A common approach to manage the spatial distribution of fishing effort to recover stocks and preserve biodiversity is through the implementation of marine reserves. These areas allow bounded populations to recover by limiting all extractive activities \citep{halpern_2002}.

The science of marine reserves has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}. Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1}. Other research has focused on the relationship between socioeconomic and governance structures and their relationship to reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, few studies simultaneously evaluate reserves from all these perspectives (*e.g.* \citet{lpezangarita_2014}).

<!-- While ecological factors like habitat representation within the MR or connectivity to other areas can determine the success of a MR, its effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed. Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs. For example, age of a reserve (*i.e.* time since its implementation) and size were key to the effectiveness of MRs in Palau \citep{friedlander_2017-oI}. In the Mediterranean, \citet{difranco_2016-Xw} identify that surveillance and enforcement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key factors that increased stock health and income to fishers. At a global level, enforcement, age, size, and isolation are important factors that determine the effectiveness of the reserves \citep{edgar_2014-UO}. However, MRs represent part of a more complex social-ecological systems governed by the interaction between humans and the environment. In this work we evaluate community-based marine reserves in Mexico using a triple bottom line approach by using ecological, socioeconomic, and governance indicators that provide a holistic view of the reserves \citep{halpern_2013}. -->

Implementation of marine reserves can be done by following one of three main approaches in terms of implementation. We describe these in the context of Mexican MRs, but argue that at least the first two apply elsewhere. Marine reserves are usually implemented and managed by a government agency, in this case the National Commission of Natural Protected Areas (*Comisión Nacional de Áreas Naturales Protegidas*, CONANP). While CONANP has made efforts to incorporate stakeholders in the design and implementation phases, these are still characterized by a heavy top-down process. A second approach is the implementation of community-based marine reserves, usually placed within areas of exclusive access (*i.e.* TURFs). Community-based spatial closures occur in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}. This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However, without legal recognition these are difficult to enforce and fishers rely on the exclusive access granted by the TURF. In an effort to bridge this normative gap, Civil Society Organizations (CSOs) served as the link between fishers and government, and set out to create a legal framework that solve this governance issue. In 2014, a new norm was created, allowing fishers to request the legal recognition of a community-based reserve under the name of "Fishing Refugia" \citep{nom}. These can be implemented as temporal or partial reserves, which can protect one, some, or all resources within them. Since then, **41** of community-based marine reserves along the Pacific, Gulf of California, and Mexican Caribbean coastlines have gained legal recognition, but no formal evaluation of their effectiveness has taken place.

This work combines causal inference techniques and the social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico. The objective of this work is twofold. First, provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries. And second, perform the first formal evaluation of Fishing Refugia in Mexico and identify areas where improvement or adjustment might result in increased effectiveness. On both cases, we draw from the lessons learned and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries.


# Materials and Methods

## Study area

We evaluate community-based marine reserves from three coastal communities located in the Pacific coast of Baja California and the Mexican Caribbean (Fig \ref{fig:map}). All communities are organized as fishing cooperatives that hold Territorial Use Rights for Fisheries (TURFs). Isla Natividad lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests and rocky reefs are the predominant habitats. The island is home to the *Buzos y Pescadores de la Baja California* fishing cooperative, whose main resource by value is the spiny lobster (*Panulirus interruptus*). Other resources like finfish, sea cucumber, red sea urchin, snail, and abalone are also important sources of income. In 2006, the community decided to implement two community-based marine reserves within their fishing grounds to protect commercially important invertebrate species (mainly lobster and abalone). These reserves obtained legal recognition in 2018, but were well enforced since their implementation in 2006.

The other two communities are Maria Elena and Punta Herrero, which are located in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems( Fig \ref{fig:map}C). Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); Punta Herrero is home to the *SCPP José María Azcorra* cooperative. Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight marine reserves in 2012, and four marine reserves in 2013, respectively. All these reserves are legally recognized as Fishing Refugia since their creation.

## Data collection

```{r load_data}
invert <- read.csv(file = here::here("data", "invertebrates.csv"),
                stringsAsFactors = T)

fish <- read.csv(file = here::here("data", "fish.csv"),
                stringsAsFactors = T)

conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
                stringsAsFactors = T)

gob <- read.csv(here::here("data" ,"governance.csv"),
                strip.white = T,
                stringsAsFactors = T)
```

We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions. Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)). Trained divers record richness and abundances of fish and invertebrate species in the reserves and control sites. Size structures are also collected during fish surveys. We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had a similar probability of being selected as reserves during the design phase. We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed annually, and at least once before implementation of the reserves. Table \ref{table:com_sum} shows a summary of the reserves included in this study.

```{r load_spatial}
# Determine a projection for the maps
proj <- "+proj=longlat +datum=NAD27"

# Load coastline for Mexico and World
data(World)
W <- spTransform(World, proj) %>% 
  st_as_sf()

# Load high-res mexican coastline
coastline_mx <- readRDS(here::here("raw_data", "spatial", "coastline_mx.rds"))

# Load ZRP shapefile
zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
  filter(Name %in% c("El Cabezo",
                     "Gallineros",
                     "Punta Loria",
                     "San Roman Norte",
                     "San Roman Sur",
                     "Anegado de Chal",
                     "Laguna Canche Balam",
                     "Niluc",
                     "Mimis",
                     "El Faro (Pesca de Langosta)",
                     "El Faro")) %>% 
  st_transform(crs = proj) %>% 
  mutate(evaluated = ifelse(Name %in% c("El Cabezo", "El Faro (Pesca de Langosta)", "El Faro"), T, F))

# Load Community reserve shapefile
rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
  filter(Nombre %in% c("La_Plana", "PuntaPrieta")) %>% 
  st_transform(crs = proj)

turf <- st_read(dsn = here::here("raw_data", "spatial"), layer = "turfs", quiet = T) %>% 
  filter(Coop %in% c("SCPP Buzos y Pescadores de la Baja California", "SCPP Cozumel", "SCPP Jose Maria Azcorra"))

```

```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A). Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the Yucatan Peninsula. Blue polygons represent the TURFs, and red polygons the marine reserves."}
# Define reference map

## Declare points for reference map
location_labels <- data.frame(X = c(-115.24, -87.4611756),
                              Y = c(27.85, 19.3183386),
                              Location = c("Isla Natividad", "Punta Herrero"),
                              Region = c("Pacific", "Caribbean"),
                              stringsAsFactors = F)

## Create map
mex <- ggplot() +
  geom_sf(data = W) +
  geom_point(data = location_labels, aes(x = X, y = Y), color = "black", fill = "steelblue", size = 3, shape = 21) +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "none",
        panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 7)) +
  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863)) +
  annotate(geom = "text", x = -102, y = 25, label = "Mexico") +
  annotate(geom = "text", x = -90, y = 25, label = "Gulf of\nMexico") +
  annotate(geom = "text", x = -112, y = 20, label = "Pacific\nOcean") +
  labs(x = "", y = "")

# Map for SAM

## Declare points for SAM
sam_points <- data.frame(x = c(-87.4646745, -87.539104),
                         y = c(19.3229767, 19.396729),
                         location = c("Punta Herrero", "Maria Elena"))

## Create map
ph <- ggplot() +
  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
  geom_sf(data = coastline_mx) +
  lims(x = c(-87.7, -87.35), y = c(19.15, 19.6)) +
  geom_sf(data = zrp, fill = "red2") +
  theme_classic() +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 7)) +
  geom_label_repel(data = sam_points, aes(x = x, y = y, label = location), nudge_y = 0.05, nudge_x = 0.05) +
  labs(x = "", y = "") +
  scalebar(data = turf[3,], dist = 5, dd2km = T, model = 'WGS84', location = "bottomright", st.size = 2, st.dist = 0.05, height = 0.1) +
  scale_x_continuous(breaks = seq(-87.7, -87.3, 0.1), limits = c(-87.7, -87.3))

# Map for PBC
pbc <- ggplot() +
  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
  lims(x = c(-115.65, -115.109), y = c(27.7, 28)) +
  geom_sf(data = coastline_mx) +
  geom_sf(data = rescom, fill = "red2") +
  theme_classic() +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 7)) +
  labs(x = "", y = "") +
  scalebar(data = turf[4,], dist = 10, dd2km = T, model = 'WGS84', location = "bottomleft", st.size = 2, st.dist = 0.05, height = 0.1)

left <- plot_grid(mex, pbc, ncol = 1, labels = "AUTO")

plot_grid(left, ph, labels = c("", "C"), ncol = 2)
```

```{r}
tibble(Community = c("Isla Natividad", "Maria Elena", "Punta Herrero"),
       TURF = c(889.5, 353.1, 299.7),
       reserve = c(1.53, 0.10, 0.43),
       percent = reserve / TURF * 100,
       implementation = c(2006, 2012, 2013)) %>%
  knitr::kable(format = "latex",
               caption = "\\label{table:com_sum} Summary of commuity--based marine reserves by community.",
               col.names = c("Community", "TURF area ($km^2$)", "Reserve area ($km^2$)", "Percent as reserves","Year of implementation"),
               escape = F)
```


Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA). Data contain monthly lobster landings (Kg) and revenues (MXP) from 2000 to 2014 for cooperatives with and without marine reserves(**Fig S1**). All cooperatives of each region (*i.e.* Pacific and Caribbean) incorporated in this analysis, belong to larger Cooperative Federations, and are exposed to the same markets and institutional frameworks \citep{mccay_2017-1m,ayer_2018}, making them plausible controls. Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted by the Consumer Price Index for Mexico \citep{oecd_2017-VV} as:

\begin{equation}
I_t = RI_t\times\frac{CPI_t}{CPI_T}
\label{eqn:cpi}
\end{equation}

Where $I_t$ represents the adjusted income for year $t$ as the product between the reported income for that year and the ratio between the consumer price index in that year ($CPI_t$) to the most recent year’s consumer price index ($CPI_T$).

Data for the qualitative analysis of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities. These include information on the resource system, the resource units, actors, and the governance system itself (**S1 Table**).

## Data analysis

We evaluate the effect that marine reserves have had on four ecological and two socioeconomic indicators (Table \ref{table:indicators}). Recall that reserves were implemented to protect lobster and other benthic invertebrates. However, we also use the available fish datato test for associated co-benefits.

```{r}
tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
           Indicador = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
       Units = c("org $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "Kg $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "M MXP", "Metric Tonnes")) %>% 
  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
  kable_styling(latex_options = "HOLD_position")
```

We use a difference-in-differences analysis to evaluate these indicators. This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments (*i.e.* reserve / control sites \citet{moland_2013-VP,Villasenor-Derbez_2018}). The analysis of ecological indicators is performed with a multiple linear regression of the form:

\begin{equation}
I_{itj} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
\label{eqn:reg_bio}
\end{equation}

Where year-fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site (See Table \ref{table:com_sum}). When evaluating biomass and densities of the entire benthic or fish communities, we include $\sigma_j$ to control for species-fixed effects.

Socioeconomic indicators are evaluated with a similar approach. Due to data constrains, we only evaluate socioeconomic data for Isla Natividad and Maria Elena. Neighboring communities are used as counterfacutals that allow us to control for unobserved time-invariants. Each "treated" community (Isla Natividad and Maria Elena) has three counterfactual communities.

\begin{equation}
I = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon
\label{eqn:soc_reg}
\end{equation}

The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jCom$ captures community-level fixed-effects. These regressions allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}. The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively). All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.


```{r model_fit}
# Invertebrates

## Lobsters
L_IN <- invert %>%
  filter(generoespecie == "Panulirus interruptus") %>% 
  abundance(com = "Isla Natividad") %>% 
  model(sfe = F, spfe = F)

L_ME <- invert %>% 
  filter(generoespecie == "Panulirus argus") %>% 
  abundance(com = "Maria Elena") %>% 
  model(sfe = F, spfe = F)

L_PH <- invert %>% 
  filter(generoespecie == "Panulirus argus") %>% 
  abundance(com = "Punta Herrero") %>% 
  model(sfe = F, spfe = F)

## Abundance
Ni_IN <- invert %>% 
  abundance(com = "Isla Natividad") %>% 
  model(sfe = F)

Ni_ME <- invert %>% 
  abundance(com = "Maria Elena") %>% 
  model(sfe = F)

Ni_PH <- invert %>% 
  abundance(com = "Punta Herrero") %>% 
  model(sfe = F)

## Abundance commercially important

# Fish
## Biomass
B_IN <- fish %>% 
  biomass(com = "Isla Natividad") %>% 
  model(sfe = F)

B_ME <- fish%>% 
  biomass(com = "Maria Elena") %>% 
  model(sfe = F)

B_PH <- fish %>% 
  biomass(com = "Punta Herrero") %>% 
  model(sfe = F)

## Abundance
N_IN <- fish %>% 
  abundance(com = "Isla Natividad") %>% 
  model(sfe = F)

N_ME <- fish %>% 
  abundance(com = "Maria Elena") %>% 
  model(sfe = F)

N_PH <- fish %>% 
  abundance(com = "Punta Herrero") %>% 
  model(sfe = F)

```

# Results

The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community. Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size $\lambda_t$). We also provide an overview of the governance settings of each community, and discuss how these might be related to the effectiveness and performance of the reserves.

## Biological

Indicators showed ambiguous responses through time for each reserve. Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time. These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ but are not significantly different from zero ($p > 0.05$). Lobster densities were only significantly positive for Isla Natividad on the sixth year (*i.e.* 2012; $p < 0.05$), a year after the hypoxia events described by \citet{micheli_2012-EU} caused mass mortality of organisms. Likewise, no changes were detected in fish biomass or invertebrate and fish densities (\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends. Full tables with model coefficients are presented in the supplementary materials (**S2 Table**, **S3 Table**, **S4 Table**).

## Socioeconomic

Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}). For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, asuggesting that these are plausible controls. However, effect sizes do not change after the implementation of the reserve. Again, the negative coefficient observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events. The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$). The three years of post-implementation data for Maria Elena do not show a significant effect of the reserve. Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities. However, these changes are not significant and are associated to increased variation. All regression coefficients for each community and indicator are presented in **S5 Table**.

## Governance

Although we have little information on the social dimension of these fisheries, we can use the social-ecological systems framework (**S1 Table**) to analyze the performance of each governance system (**S6 Table**). Our analysis shows that all of the systems analyzed share similarities in their Governance system which is based on cooperatives (GS5.2.3.2), with strong rules in use that include Operational rules (GS6.2), Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even Territorial use communal rights (GS6.1.4.3). However, we identified important differences in terms of the actors, resource systems, and resource units. Although all communities show a high level of leadership (A5), the level of trust (A6.1) is lower in Punta Herrero. In general, the presence and success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources they depend upon \citep{jupiter_2017}. The enabling conditions for conservation seem to be strongly present in all communities. Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperative. These conditions have favored the development of an efficient community-based enforcement systems.

\clearpage

```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for lobster densities (*Panulirus spp*; A), fish biomass (B), invertebrate densities (C), and fish densities (D). Plots are ordered by survey type (left column: invertebrates; right column: fish). Points are jittered hotizontally to avoid overplotting. Points indicate the effect size, and errorbars standard errors. Years have been centered to year of implementation."}

# Invert plots
## Lobsters
L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
                     model_names = c("IS", "ME", "PH"))

## Abundance
I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
                  model_names = c("IN", "ME", "PH"))

# Fish plots
## Biomass
F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
                 model_names = c("IS", "ME", "PH"))

## Abundance
F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
                  model_names = c("IN", "ME", "PH")) +
  theme(legend.justification = c(1, 0.1),
        legend.position = c(1, 0),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.direction = "horizontal")

cowplot::plot_grid(plotlist = list(L_plot, F_B, I_N, F_N),
                   labels = "AUTO",
                   ncol = 2)
```

```{r, results = "asis"}
C_IN <- conapesca %>% 
  filter(years_centered > -10) %>% 
  mutate(years_centered = as.factor(years_centered),
         years_centered = fct_relevel(years_centered, "0")) %>% 
  filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
  filter(zona == "PN") %>%
  lm(pesovivo/1e3 ~ grupo * years_centered + unidadeconomica, data = .)

C_ME <- conapesca %>% 
  filter(years_centered > -10) %>% 
  mutate(years_centered = as.factor(years_centered),
         years_centered = fct_relevel(years_centered, "0")) %>% 
  filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
  filter(zona == "MC") %>%
  lm(pesovivo/1e3 ~ grupo * years_centered + unidadeconomica, data = .)

V_IN <- conapesca %>% 
  filter(years_centered > -10) %>% 
  mutate(years_centered = as.factor(years_centered),
         years_centered = fct_relevel(years_centered, "0")) %>% 
  filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
  filter(zona == "PN") %>%
  lm(valor/1e6 ~ grupo * years_centered + unidadeconomica, data = .)

V_ME <- conapesca %>% 
  filter(years_centered > -10) %>% 
  mutate(years_centered = as.factor(years_centered),
         years_centered = fct_relevel(years_centered, "0")) %>% 
  filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
  filter(zona == "MC") %>%
  lm(valor/1e6 ~ grupo * years_centered + unidadeconomica, data = .)

#soc_res se fue

plot_socioeco <- function(model_list, model_names, ylab){
  
  extract_filter <- function(model, name){
    model %>% 
      robust_se() %>% 
      filter(grepl(pattern = "grupoTreated:years_centered", x = term)) %>% 
      mutate(years_centered = as.numeric(gsub(pattern = "grupoTreated:years_centered",
                                              replacement = "",
                                              x = term)),
             Community = name)
  }
  
  purrr::map2_df(model_list, model_names, extract_filter) %>% 
    ggplot(aes(x = years_centered, y = estimate, group = Community)) +
    geom_hline(yintercept = 0,
               linetype = "dashed",
               size = 1) +
    geom_vline(xintercept = 0,
               linetype = "dashed",
               size = 1) +
    geom_errorbar(aes(color = Community,
                      ymin = estimate-std.error,
                      ymax = estimate + std.error),
                  width = 0.1,
                  size = 1,
                  position = position_dodge(width = 0.4)) +
    geom_point(aes(color = Community, shape = Community),
               size = 4,
               position = position_dodge(width = 0.4),
               alpha = 0.5) +
    scale_color_brewer(palette = "Set1") +
    labs(x = "Years since implementation", y = ylab) +
    theme(text = element_text(size = 12))
}
```


```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles)"}

c_plot <- plot_socioeco(model_list = list(C_IN, C_ME),
                        model_names = list("IN", "ME"),
                        ylab = quo(lambda[t]))

v_plot <- plot_socioeco(model_list = list(V_IN, V_ME),
                        model_names = list("IN", "ME"),
                        ylab = quo(lambda[t]))

plot_grid(c_plot, v_plot, ncol = 1, labels = "AUTO")

```


\clearpage

# Discussion

Our results indicate that marine reserves do not achieve the main objective of increasing lobster densities. No co-benefits were identified when using other ecological indicators other than the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \citep{micheli_2012-EU}. The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation. The lack of expected effectiveness poses the question: why do these communities continue to support the reserves? Understanding the social-ecological context in which these communities and their reserves operate might provide insights to this question. Here we touch on potential shortcomings in our analysis, and discuss our findings in the context of existing literature and our social-ecological system analysis. Finally, we provide recommendations to guide the implementation of new (and improve effectiveness of) community-based marine reserves in small-scale fisheries.

Some works evaluate marine reserves by performing inside-outside \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or before-after comparisons \citep{betti_2017-lq}. The first approach does not address temporal variability, and the second can not distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}. Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness. However, this method assumes control sites are a plausible counterfactual for treated sites. This supposed that treated sites would have followed the same trend as control sites, had the reserves not been implemented. Nonetheless, overall trends for each site don't show any significant increases, supporting our findings of lack of change in the indicators used (**S2 Figure**, **S3 Figure**, **S4 Figure**, **S5 Figure**, **S6 Figure**).

Literature shows that age, isolation, and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO}. Isla Natividad has the oldest reserve, and all communities are fairly isolated, and have a well-established community-based enforcement system. With these characteristics, one would expect the reserves to be effective. However, another key condition for effectiveness is reserve is size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to reserves being too small. Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sesile invertebrates, but not lobsters. Maria Elena and Punta Herrero are relatively young reserves, and it is known that community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}. Work by \citet{ayer_2018} shows these communities support the implementation of marine reserves. Nevertheless, fishers may favor implementation of reserves that pose low fishing costs either because of their location or size. The economic data support our claim that reserves are small, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes, there are other plausible explanations for the observed lack of effectiveness. Marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2006}. However, both lobster fisheries were, at some point, certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \cite{dof_website_1993}. It is uncertain whether that such a well-managed fishery will experience additional benefits from marine reserves.

While reserves fail to provide fishery benefits, there are a number of additional ecological, fisheries, and social benefits. Marine reserves provide protection to a wider range of species and vulnerable habitat, like coral reefs. These sites can serve as an insurance against environmental shocks or mistakes in fisheries management \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}. Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital. Furthermore, showing commitment to marine conservation allows fishers to have greater bargaining power and leverage over fisheries management.

Community-based marine reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented. Lessons learned from these cases can guide implementation of community-based marine reserves elsewhere. For the particular case of the marine reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities. At the broader scale, having full community support surely represents an advantage, but it is important for marine reserves to meet essential design principles such as size and placement. Community-based marine reserves might have more benefits that result from indirect effects of the reserves, which should be taken into account when evaluating the outcomes of similar projects.

\section*{Conflict of Interest Statement}

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}

JC and EA analyzed and interpreted data, discussed the results, and wrote the first draft. AS, SF and JT discussed the results and edited the manuscript.

\section*{Funding}

JCVD CONACyT (Beca de Posgrados en el extranjero, CVU 669403) + LAFF + El proyecto de COBI que haya pagado esto

**ASC**

**SF**

**JT**

**Que fuentes de financiamiento incluyo par ustedes?**

\section*{Acknowledgments}
The authors wish to acknowledge Arturo Hernández and Imelda Amador for contributions on the governance data, as well as pre-processing biological data. This study would have not been possible without the effort by members of the communities here mentioned, who collected the biological data.

\section*{Supplemental Data}
 \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder
 
\paragraph*{S1 Figure}
\label{S1_Figure}
Map of control and treated sites in A and control and treated landings in B

\paragraph*{S2 Figure}
\label{S2_Figure}
Time series of biological indicators for IN

\paragraph*{S3 Figure}
\label{S3_Figure}
Time series of biological indicators for  ME

\paragraph*{S4 Figure}
\label{S4_Figure}
Time series of biological indicators for  PH

\paragraph*{S5 Figure}
\label{S5_Figure}
Time series of economic indicators for ME

\paragraph*{S6 Figure}
\label{S6_Figure}
Time series of economic indicators for PH

\paragraph*{S1 Table}
\label{S1_Table}
Coefficient estimates for biological indicators in Isla Natividad

\paragraph*{S2 Table}
\label{S2_Table}
Coefficient estimates for biological indicators in Maria Elena

\paragraph*{S3 Table}
\label{S3_Table}
Coefficient estimates for biological indicators in Punta Herrero

\paragraph*{S4 Table}
\label{S4_Table}
Coefficient estimates for economic indicators

```{r, results = "asis", eval = F}

se <- list(robust_se_list(L_IN),
           robust_se_list(L_ME),
           robust_se_list(L_PH))

stargazer::stargazer(list(L_IN, L_ME, L_PH),
                     se = se,
                     t.auto = T,
                     p.auto = F,
                     type = "html",
                     single.row = T,
                     model.names = F,
                     column.labels = c("IN", "ME", "PH"),
                     dep.var.labels.include = FALSE,
                     model.numbers = FALSE,
                     digits = 2,
                     intercept.bottom = F,
                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
                     out = "lobster_models.html")
```


```{r, eval = F}
se <- list(robust_se_list(B_IN),
           robust_se_list(B_ME),
           robust_se_list(B_PH))

stargazer::stargazer(list(B_IN, B_ME, B_PH),
                     se = se,
                     t.auto = T,
                     p.auto = F,
                     type = "html",
                     single.row = T,
                     model.names = F,
                     column.labels = c("IN", "ME", "PH"),
                     dep.var.labels.include = FALSE,
                     model.numbers = FALSE,
                     digits = 2,
                     intercept.bottom = F,
                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
                     out = "lobster_models.html")
```

```{r, eval = F}
se <- list(robust_se_list(Ni_IN),
           robust_se_list(Ni_ME),
           robust_se_list(Ni_PH))

stargazer::stargazer(list(Ni_IN, Ni_ME, Ni_PH),
                     se = se,
                     t.auto = T,
                     p.auto = F,
                     type = "html",
                     single.row = T,
                     model.names = F,
                     column.labels = c("IN", "ME", "PH"),
                     dep.var.labels.include = FALSE,
                     model.numbers = FALSE,
                     digits = 2,
                     intercept.bottom = F,
                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
                     out = "lobster_models.html")
```

```{r, eval = F}
se <- list(robust_se_list(N_IN),
           robust_se_list(N_ME),
           robust_se_list(N_PH))

stargazer::stargazer(list(N_IN, N_ME, N_PH),
                     se = se,
                     t.auto = T,
                     p.auto = F,
                     type = "html",
                     single.row = T,
                     model.names = F,
                     column.labels = c("IN", "ME", "PH"),
                     dep.var.labels.include = FALSE,
                     model.numbers = FALSE,
                     digits = 2,
                     intercept.bottom = F,
                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
                     out = "lobster_models.html")
```

```{r, eval = F}
se <- list(robust_se_list(C_IN),
           robust_se_list(C_ME),
           robust_se_list(V_IN),
           robust_se_list(V_ME))

stargazer::stargazer(list(C_IN, C_ME, V_IN, V_ME),
                     se = se,
                     t.auto = T,
                     p.auto = F,
                     type = "html",
                     single.row = T,
                     model.names = F,
                     column.labels = c("Landings IN", "Landings ME", "Revenue IN", "Revenue ME"),
                     dep.var.labels.include = FALSE,
                     model.numbers = FALSE,
                     digits = 2,
                     intercept.bottom = F,
                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
                     out = "lobster_models.html")
```


\bibliographystyle{frontiersinSCNS_ENG_HUMS}
\bibliography{references}

\section*{Figure captions}

